\begin{longtable}{lcccccc}
\caption[Summary of Meta-Information of All the Models.]{\textbf{Summary of Meta-Information of All the Models.}\\The number of variables, parameters, reactions, and derived variables and parameters, and total sum are shown for each model. These values have been extracted directly from the model implementations. A derived quantity, is a quantity that is calculated inside the model. The separation between derived variable and parameter is based on expertise and is not based on a strict rule. However, a main aspect that was considered was with what quantity it was derived from. If the quantity was derived from even a single time-dependent quantity, it is considered a derived variable. The model used are Bellasio2019~\cite{bellasioGeneralisedDynamicModel2019}, Fuente2024~\cite{fuenteMathematicalModelSimulate2024}, Li2021~\cite{liImpactIonFluxes2021}, Matuszynska2016~\cite{matuszynskaMathematicalModelNonphotochemical2016}, and Saadat2021~\cite{saadatComputationalAnalysisAlternative2021}} \label{tab:models-meta} \\
\toprule
 & Variables & Parameters & Reactions & Derived Variables & Derived Parameters & Total \\
\midrule
\endfirsthead
\caption[]{\textbf{Summary of Meta-Information of All the Models.}\\The number of variables, parameters, reactions, and derived variables and parameters, and total sum are shown for each model. These values have been extracted directly from the model implementations. A derived quantity, is a quantity that is calculated inside the model. The separation between derived variable and parameter is based on expertise and is not based on a strict rule. However, a main aspect that was considered was with what quantity it was derived from. If the quantity was derived from even a single time-dependent quantity, it is considered a derived variable. The model used are Bellasio2019~\cite{bellasioGeneralisedDynamicModel2019}, Fuente2024~\cite{fuenteMathematicalModelSimulate2024}, Li2021~\cite{liImpactIonFluxes2021}, Matuszynska2016~\cite{matuszynskaMathematicalModelNonphotochemical2016}, and Saadat2021~\cite{saadatComputationalAnalysisAlternative2021}} \\
\toprule
 & Variables & Parameters & Reactions & Derived Variables & Derived Parameters & Total \\
\midrule
\endhead
\midrule
\multicolumn{7}{r}{Continued on next page} \\
\midrule
\endfoot
\bottomrule
\endlastfoot
Bellasio2019 & 13 & 82 & 17 & 6 & 16 & 134 \\
Fuente2024 & 5 & 31 & 10 & 9 & 2 & 57 \\
Li2021 & 13 & 43 & 20 & 13 & 7 & 96 \\
Matuszynska2016 & 6 & 42 & 8 & 11 & 6 & 73 \\
Saadat2021 & 30 & 167 & 46 & 25 & 24 & 292 \\
\end{longtable}
